---
title: "pfcsp mosquito abdomen sharing model prep descriptives"
author: "Kelsey Sumner"
date: "10/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up work environment

## Load packages

```{r}
library(readr)
library(lme4)
library(moments)
library(ggplot2)
library(epiDisplay)
library(dplyr)
```

## Read in the csp abdomen data set

This data set shows the pairs of human asymptomatic and symptomatic infections with mosquito abdomens that were Pf positive and collected within the same household as the human, 0-14 days after the human infection was reported.

```{r}
csp_abdomens <- read_rds("~/Desktop/clean_ids_haplotype_results/CSP/spat21_csp_edgelist_abdomen_29OCT2019.rds")

# load in the data set (the haplotypes after chimeras have been removed and haplotypes censored - seqtab_final.rds)
csp_haplotypes <- read_rds("Desktop/clean_ids_haplotype_results/CSP/spat21_CSP_haplotype_table_censored_final_version_with_moi_and_ids_CLEANVERSION_30SEPT2019.rds")

```

# Check data set up prior to running models

## Merge human MOI into the data set

```{r}
# subset th csp_haplotpes data set to sample id and moi
csp_haplotypes = csp_haplotypes %>% 
  select(sample_name_dbs,haplotype_number) %>%
  rename("sample_id_human" = "sample_name_dbs","moi"="haplotype_number")

# merge the csp_abdomens and csp_haplotypes data sets to get moi
csp_abdomens = left_join(csp_abdomens,csp_haplotypes,by="sample_id_human")
length(which(is.na(csp_abdomens$moi)))
str(csp_abdomens$moi)
```

## Make sure variables all coded correctly prior to running models

Check that the exposure, outcome, and covariates are coded correctly

```{r}
# make sure the exposure and outcome are coded correctly
str(csp_abdomens$aim2_exposure)
csp_abdomens$aim2_exposure = as.factor(csp_abdomens$aim2_exposure)
str(csp_abdomens$haps_shared)
str(csp_abdomens$village_name)
csp_abdomens$village_name = as.factor(csp_abdomens$village_name)
str(csp_abdomens$pfr364Q_std_combined)
str(csp_abdomens$age_cat_baseline)
csp_abdomens$age_cat_baseline = as.factor(csp_abdomens$age_cat_baseline)
str(csp_abdomens$sample_id_human)
str(csp_abdomens$unq_memID)
str(csp_abdomens$HH_ID)
str(csp_abdomens$age_all_baseline)
csp_abdomens$age_all_baseline = as.numeric(csp_abdomens$age_all_baseline)
```

Look at a summary of the variables in the analysis set. 

```{r}
# look at a summary of aim2_exposure
table(csp_abdomens$aim2_exposure, useNA = "always")
# look at haps_shared
summary(csp_abdomens$haps_shared)
sd(csp_abdomens$haps_shared)
skewness(csp_abdomens$haps_shared)
kurtosis(csp_abdomens$haps_shared)
# look at pfr364Q_combined_std
summary(csp_abdomens$pfr364Q_std_combined)
sd(csp_abdomens$pfr364Q_std_combined)
skewness(csp_abdomens$pfr364Q_std_combined)
kurtosis(csp_abdomens$pfr364Q_std_combined)
# look at age_cat
table(csp_abdomens$age_cat_baseline, useNA = "always")
table(csp_abdomens$age_all_baseline, useNA = "always")
summary(csp_abdomens$age_all_baseline)
table(csp_abdomens$age_cat_baseline,csp_abdomens$age_all_baseline,useNA = "always")
# look at village_name
table(csp_abdomens$village_name, useNA = "always")
# look at sample_id_human
table(csp_abdomens$sample_id_human, useNA = "always")
table(csp_abdomens$sample_id_abdomen, useNA = "always")
# look at unq_memID
table(csp_abdomens$unq_memID, useNA = "always")
# look at HH_ID
table(csp_abdomens$HH_ID, useNA = "always")
# note: looks like a lot of multiple pairings where have a mosquito with many people and many people with a mosquito, especially in K01
# might want to do weighting to account for this
```

# Remove Sitabicha from analysis

Sitabicha was only included in 12 pairs (3%) of the final data set, so removing from the analysis because the small numbers are not representative of Sitabicha as a whole and will cause sparse data and model convergence issues. 

```{r}
csp_abdomens = csp_abdomens %>%
  filter(!(village_name=="Sitabicha"))
```


Look at a summary of the variables in the analysis set with Sitabicha removed. 

```{r}
# look at a summary of aim2_exposure
table(csp_abdomens$aim2_exposure, useNA = "always")
# look at haps_shared
summary(csp_abdomens$haps_shared)
sd(csp_abdomens$haps_shared)
skewness(csp_abdomens$haps_shared)
kurtosis(csp_abdomens$haps_shared)
# look at pfr364Q_combined_std
summary(csp_abdomens$pfr364Q_std_combined)
sd(csp_abdomens$pfr364Q_std_combined)
skewness(csp_abdomens$pfr364Q_std_combined)
kurtosis(csp_abdomens$pfr364Q_std_combined)
# look at age_cat
table(csp_abdomens$age_cat_baseline, useNA = "always")
table(csp_abdomens$age_all_baseline, useNA = "always")
summary(csp_abdomens$age_all_baseline)
table(csp_abdomens$age_cat_baseline,csp_abdomens$age_all_baseline,useNA = "always")
# look at village_name
table(csp_abdomens$village_name, useNA = "always")
# look at sample_id_human
table(csp_abdomens$sample_id_human, useNA = "always")
table(csp_abdomens$sample_id_abdomen, useNA = "always")
# look at unq_memID
table(csp_abdomens$unq_memID, useNA = "always")
# look at HH_ID
table(csp_abdomens$HH_ID, useNA = "always")
# note: looks like a lot of multiple pairings where have a mosquito with many people and many people with a mosquito, especially in K01
# might want to do weighting to account for this
```

## Assess functional form coding for pfr364Q_combined_std

Assessed functional form coding for the variable pfr364Q_std_combined, which measured the P. falciparum parasites/uL in the human blood samples. 

First plot the natural form using a LOWESS curve. 

```{r}
# look at the natural form of pfr364Q_std_combined through the LOWESS curve
# plot the lowess graph 
plot_loess = ggplot(data=csp_abdomens) +
  geom_point(aes(x=pfr364Q_std_combined,y=haps_shared)) +
  geom_smooth(aes(x=pfr364Q_std_combined,y=haps_shared),method="loess")
plot_loess
```


Now look at the linear coding of pfr364Q_std_combined.

```{r}
model1=glm(haps_shared~pfr364Q_std_combined,family=poisson(link = "log"),data=csp_abdomens) 
summary(model1) 
# exponentiate coefficient
exp(-8.831e-06)
# pull out log likelihood test value
logLik(model1) 
# plot the linear model fit
linear_plot = ggplot(csp_abdomens, aes(x=pfr364Q_std_combined, y=haps_shared)) + 
  labs(title="LINEAR Model",
       x="Parasite density",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Linear is red. 95%CI is for the linear trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
linear_plot
```

Now look at the quadratic coding of pfr364Q_std_combined.

```{r}
# first make the quadratic term
csp_abdomens$pfr364Q_std_combined_quad = csp_abdomens$pfr364Q_std_combined*csp_abdomens$pfr364Q_std_combined
# then run the model
model2=glm(haps_shared~pfr364Q_std_combined+pfr364Q_std_combined_quad,family=poisson(link = "log"),data=csp_abdomens) 
summary(model2) 
# exponentiate coefficient
exp(-1.321e-04)
exp(2.277e-09)
# pull out log likelihood test value
logLik(model2) 
csp_abdomens$pred.quad = model2$fitted.values
# plot the model fit
quadratic_plot = ggplot(csp_abdomens, aes(x=pfr364Q_std_combined, y=haps_shared)) + 
  labs(title="QUADRATIC Model",
       x="Parasite density",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Quadratic is red [y~ (x+x^2)]. 95%CI is for the quadratic trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~poly(x,2), color="red", linetype="dashed") +
  theme_bw()
quadratic_plot
```

Now look at the cubic coding of pfr364Q_std_combined.

```{r}
# first make the cubic term
csp_abdomens$pfr364Q_std_combined_cubed = csp_abdomens$pfr364Q_std_combined*csp_abdomens$pfr364Q_std_combined*csp_abdomens$pfr364Q_std_combined
# then run the model
model3=glm(haps_shared~pfr364Q_std_combined+pfr364Q_std_combined_quad+pfr364Q_std_combined_cubed,family=poisson(link = "log"),data=csp_abdomens) 
summary(model3) 
# exponentiate coefficient
exp(-1.317e-04)
exp(2.213e-09)
exp(1.015e-15)
# pull out log likelihood test value
logLik(model3) 
csp_abdomens$pred.quad.cubed = model3$fitted.values
# plot the model fit
cubic_plot = ggplot(csp_abdomens, aes(x=pfr364Q_std_combined, y=haps_shared)) + 
  labs(title="CUBIC Model",
       x="Parasite density",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Cubic is red [y~ (x+x^2+x^3)]. 95%CI is for the cubic trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~poly(x,3), color="red", linetype="dashed") +
  theme_bw()
cubic_plot
```


Now look at the categorical coding of pfr364Q_std_combined.

```{r}
# first make the categorical term
csp_abdomens$pfr364Q_std_combined_cat = ifelse(csp_abdomens$pfr364Q_std_combined<100,"under_rdt_detection",
  "over_rdt_detection")
table(csp_abdomens$pfr364Q_std_combined_cat,csp_abdomens$pfr364Q_std_combined, useNA = "always")
csp_abdomens$pfr364Q_std_combined_cat = as.factor(csp_abdomens$pfr364Q_std_combined_cat)
# then run the model
model4=glm(haps_shared~pfr364Q_std_combined_cat,family=poisson(link = "log"),data=csp_abdomens) 
summary(model4) 
# exponentiate coefficient
exp(0.89882)
# pull out log likelihood test value
logLik(model4) 
csp_abdomens$pred.cat = model4$fitted.values
# plot the model fit
categorical_plot = ggplot(csp_abdomens, aes(x=pfr364Q_std_combined, y=pred.cat)) + 
  labs(title="CATEGORICAL Model",
       x="Parasite density",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Categorical is red. 95%CI is for the categorical trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
categorical_plot
```

Now look at the natural log (ln) coding of pfr364Q_std_combined.

```{r}
# first make the categorical term
csp_abdomens$pfr364Q_std_combined_ln = log(csp_abdomens$pfr364Q_std_combined)
# then run the model
model5=glm(haps_shared~pfr364Q_std_combined_ln,family=poisson(link = "log"),data=csp_abdomens) 
summary(model5) 
# exponentiate coefficient
exp(-0.12288)
# pull out log likelihood test value
logLik(model5) 
# plot the model fit
ln_plot = ggplot(csp_abdomens, aes(x=pfr364Q_std_combined_ln, y=pred.cat)) + 
  labs(title="NATURAL LOG Model",
       x="Parasite density",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Ln is red. 95%CI is for the ln trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
ln_plot
```





Will do log-likelihood test to compare nested models for differential coding of the total number of mosquitoes in the households.

```{r}
# compare quadratic and linear models
lrtest(model1,model2) # quadratic is better than linear
# compare cubic and linear models
lrtest(model1,model3) # cubic is better than linear
# compare the quadratic and cubic
lrtest(model2,model3) # quadratic is better than cubic
```





## Assess functional form coding for age

Assessed functional form coding for the age of the participants at baseline.

First plot the natural form using a LOWESS curve. 

```{r}
# look at the natural form of age through the LOWESS curve
# plot the lowess graph 
plot_loess = ggplot(data=csp_abdomens) +
  geom_point(aes(x=age_all_baseline,y=haps_shared)) +
  geom_smooth(aes(x=age_all_baseline,y=haps_shared),method="loess")
plot_loess
```

Now look at the linear coding of age_all_baseline.

```{r}
model1=glm(haps_shared~age_all_baseline,family=poisson(link = "log"),data=csp_abdomens) 
summary(model1) 
# exponentiate coefficient
exp(0.003648)
# pull out log likelihood test value
logLik(model1) 
# plot the linear model fit
linear_plot = ggplot(csp_abdomens, aes(x=age_all_baseline, y=haps_shared)) + 
  labs(title="LINEAR Model",
       x="Age",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Linear is red. 95%CI is for the linear trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
linear_plot
```

Now look at the quadratic coding of age_all_baseline.

```{r}
# first make the quadratic term
csp_abdomens$age_all_baseline_quad = csp_abdomens$age_all_baseline*csp_abdomens$age_all_baseline
# then run the model
model2=glm(haps_shared~age_all_baseline+age_all_baseline_quad,family=poisson(link = "log"),data=csp_abdomens) 
summary(model2) 
# exponentiate coefficient
exp(0.0289608)
exp(-0.0003769)
# pull out log likelihood test value
logLik(model2) 
csp_abdomens$pred.quad = model2$fitted.values
# plot the model fit
quadratic_plot = ggplot(csp_abdomens, aes(x=age_all_baseline, y=haps_shared)) + 
  labs(title="QUADRATIC Model",
       x="age",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Quadratic is red [y~ (x+x^2)]. 95%CI is for the quadratic trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~poly(x,2), color="red", linetype="dashed") +
  theme_bw()
quadratic_plot
```


Now look at the cubic coding of age_all_baseline.

```{r}
# first make the cubic term
csp_abdomens$age_all_baseline_cubed = csp_abdomens$age_all_baseline*csp_abdomens$age_all_baseline*csp_abdomens$age_all_baseline
# then run the model
model3=glm(haps_shared~age_all_baseline+age_all_baseline_quad+age_all_baseline_cubed,family=poisson(link = "log"),data=csp_abdomens) 
summary(model3) 
# exponentiate coefficient
exp(-3.783e-02)
exp(2.251e-03)
exp(-2.555e-05)
# pull out log likelihood test value
logLik(model3) 
csp_abdomens$pred.cubed = model3$fitted.values
# plot the model fit
cubic_plot = ggplot(csp_abdomens, aes(x=age_all_baseline, y=haps_shared)) + 
  labs(title="CUBIC Model",
       x="age",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Cubic is red [y~ (x+x^2+x^3)]. 95%CI is for the cubic trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~poly(x,3), color="red", linetype="dashed") +
  theme_bw()
cubic_plot
```

Now look at the categorical coding of pfr364Q_std_combined.

```{r}
# first make the categorical term
csp_abdomens$age_cat_baseline = factor(csp_abdomens$age_cat_baseline)
csp_abdomens$age_cat_baseline = relevel(csp_abdomens$age_cat_baseline,ref = ">15 years")
# then run the model
model4=glm(haps_shared~age_cat_baseline,family=poisson(link = "log"),data=csp_abdomens) 
summary(model4) 
# exponentiate coefficient
exp(-0.24660)
exp(-0.31650)
# pull out log likelihood test value
logLik(model4) 
csp_abdomens$pred.cat = model4$fitted.values
# plot the model fit
categorical_plot = ggplot(csp_abdomens, aes(x=age_cat_baseline, y=pred.cat)) + 
  labs(title="CATEGORICAL Model",
       x="age",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Categorical is red. 95%CI is for the categorical trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
categorical_plot
```


Now look at the natural log (ln) coding of age_all_baseline.

```{r}
# first make the categorical term
csp_abdomens$age_all_baseline_ln = log(csp_abdomens$age_all_baseline)
# then run the model
model5=glm(haps_shared~age_all_baseline_ln,family=poisson(link = "log"),data=csp_abdomens) 
summary(model5) 
# exponentiate coefficient
exp(0.08881)
# pull out log likelihood test value
logLik(model5) 
# plot the model fit
ln_plot = ggplot(csp_abdomens, aes(x=pfr364Q_std_combined_ln, y=pred.cat)) + 
  labs(title="NATURAL LOG Model",
       x="Parasite density",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Ln is red. 95%CI is for the ln trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
ln_plot
```



Will do log-likelihood test to compare nested models for differential coding of the total number of mosquitoes in the households.

```{r}
# compare quadratic and linear models
lrtest(model1,model2) # quadratic is better than linear
# compare cubic and linear models
lrtest(model1,model3) # cubic is better than linear
# compare the quadratic and cubic
lrtest(model2,model3) # cubic is better than quadratic
```







## Assess functional form coding for moi

Assessed functional form coding for the moi of the participants at baseline.

First plot the natural form using a LOWESS curve. 

```{r}
# look at the natural form of age through the LOWESS curve
# plot the lowess graph 
plot_loess = ggplot(data=csp_abdomens) +
  geom_point(aes(x=moi,y=haps_shared)) +
  geom_smooth(aes(x=moi,y=haps_shared),method="loess")
plot_loess
```

Now look at the linear coding of moi.

```{r}
model1=glm(haps_shared~moi,family=poisson(link = "log"),data=csp_abdomens) 
summary(model1) 
# exponentiate coefficient
exp(0.17340)
# pull out log likelihood test value
logLik(model1) 
# plot the linear model fit
linear_plot = ggplot(csp_abdomens, aes(x=moi, y=haps_shared)) + 
  labs(title="LINEAR Model",
       x="MOI",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Linear is red. 95%CI is for the linear trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
linear_plot
```

Now look at the quadratic coding of moi.

```{r}
# first make the quadratic term
csp_abdomens$moi_quad = csp_abdomens$moi*csp_abdomens$moi
# then run the model
model2=glm(haps_shared~moi+moi_quad,family=poisson(link = "log"),data=csp_abdomens) 
summary(model2) 
# exponentiate coefficient
exp(0.377701)
exp(-0.015523)
# pull out log likelihood test value
logLik(model2) 
# plot the model fit
quadratic_plot = ggplot(csp_abdomens, aes(x=moi, y=haps_shared)) + 
  labs(title="QUADRATIC Model",
       x="moi",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Quadratic is red [y~ (x+x^2)]. 95%CI is for the quadratic trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~poly(x,2), color="red", linetype="dashed") +
  theme_bw()
quadratic_plot
```

Now look at the cubic coding of moi.

```{r}
# first make the cubic term
csp_abdomens$moi_cubed = csp_abdomens$moi*csp_abdomens$moi*csp_abdomens$moi
# then run the model
model3=glm(haps_shared~moi+moi_quad+moi_cubed,family=poisson(link = "log"),data=csp_abdomens) 
summary(model3) 
# exponentiate coefficient
exp(0.526914)
exp(-0.041696)
exp(0.001296)
# pull out log likelihood test value
logLik(model3) 
# plot the model fit
cubic_plot = ggplot(csp_abdomens, aes(x=moi, y=haps_shared)) + 
  labs(title="CUBIC Model",
       x="moi",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Cubic is red [y~ (x+x^2+x^3)]. 95%CI is for the cubic trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~poly(x,3), color="red", linetype="dashed") +
  theme_bw()
cubic_plot
```


Now look at the categorical coding of moi.

```{r}
# first make the categorical term
csp_abdomens$moi_cat = ifelse(csp_abdomens$moi < 3,"1-2 haplotypes",ifelse(
  csp_abdomens$moi > 2 & csp_abdomens$moi <8,"3-7 haplotypes","8 or more haplotypes"
))
table(csp_abdomens$moi,csp_abdomens$moi_cat,useNA = "always")
# then run the model
model4=glm(haps_shared~moi_cat,family=poisson(link = "log"),data=csp_abdomens) 
summary(model4) 
# exponentiate coefficient
exp(0.7286)
exp(1.5348)
# pull out log likelihood test value
logLik(model4) 
csp_abdomens$pred.cat = model4$fitted.values
# plot the model fit
categorical_plot = ggplot(csp_abdomens, aes(x=moi, y=pred.cat)) + 
  labs(title="CATEGORICAL Model",
       x="moi",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Categorical is red. 95%CI is for the categorical trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
categorical_plot
```

Now look at the natural log (ln) coding of moi.

```{r}
# first make the categorical term
csp_abdomens$moi_ln = log(csp_abdomens$moi)
# then run the model
model5=glm(haps_shared~moi_ln,family=poisson(link = "log"),data=csp_abdomens) 
summary(model5) 
# exponentiate coefficient
exp(0.89247)
# pull out log likelihood test value
logLik(model5) 
# plot the model fit
ln_plot = ggplot(csp_abdomens, aes(x=moi_ln, y=pred.cat)) + 
  labs(title="NATURAL LOG Model",
       x="moi",
       y="Number of haplotypes shared",
       caption="NOTE: LOESS is blue. Ln is red. 95%CI is for the ln trend alone.") +
  geom_smooth(method="loess", se = F) +                   
  geom_smooth(method="glm", formula = y ~x, color="red", linetype="dashed") +
  theme_bw()
ln_plot
```



Will do log-likelihood test to compare nested models for differential coding of the human moi.

```{r}
# compare quadratic and linear models
lrtest(model1,model2) # quadratic is better than linear
# compare cubic and linear models
lrtest(model1,model3) # cubic is better than linear
# compare the quadratic and cubic
lrtest(model2,model3) # quadratic is better than cubic
```






## Clean up the variables in the data set

```{r}
colnames(csp_abdomens)
csp_abdomens = csp_abdomens %>%
  dplyr::select(-c(pfr364Q_std_combined_quad,pfr364Q_std_combined_cubed,pfr364Q_std_combined_cubed,pfr364Q_std_combined_ln,pred.quad,pred.quad.cubed,pred.cat,pred.cubed,age_all_baseline_ln))
colnames(csp_abdomens)
```


## Relevel the village_name variables
```{r}
table(csp_abdomens$village_name, useNA = "always")
csp_abdomens$village_name = as.factor(ifelse(csp_abdomens$village_name=="Kinesamo","Kinesamo","Maruti"))
table(csp_abdomens$village_name, useNA = "always")
```


## Check for muticollinearity 

```{r}
# comparing aim2_exposure vs. haps_shared
asymptomatic = csp_abdomens[which(csp_abdomens$aim2_exposure=="asymptomatic infection"),]
symptomatic = csp_abdomens[which(csp_abdomens$aim2_exposure=="symptomatic infection"),]
asymptomatic_mean = mean(asymptomatic$haps_shared, na.rm = TRUE) 
symptomatic_mean = mean(symptomatic$haps_shared, na.rm = TRUE) 
pooled_sd = sd(append(asymptomatic$haps_shared, symptomatic$haps_shared), na.rm = TRUE)
(symptomatic_mean - asymptomatic_mean)/pooled_sd

# comparing aim2_exposure vs. age_cat_baseline
table(csp_abdomens$aim2_exposure,csp_abdomens$age_cat_baseline, useNA = "always")
firstmodel = glm(aim2_exposure ~ age_cat_baseline, data=csp_abdomens, family=binomial(link="logit"))
summary(firstmodel)
exp(-0.6628)
exp(0.8177)

# comparing aim2_exposure vs. village_name
table(csp_abdomens$aim2_exposure,csp_abdomens$village_name, useNA = "always")
firstmodel = glm(village_name ~ aim2_exposure, data=csp_abdomens, family=binomial(link="logit"))
summary(firstmodel)
exp(-1.0526)

# comparing aim2_exposure vs. pfr364Q_std_combined_cat
table(csp_abdomens$aim2_exposure,csp_abdomens$pfr364Q_std_combined_cat, useNA = "always")
firstmodel = glm(pfr364Q_std_combined_cat ~ aim2_exposure, data=csp_abdomens, family=binomial(link="logit"))
summary(firstmodel)
exp(-0.8974)

# comparing haps_shared vs. age_cat_baseline
cat1 = csp_abdomens[which(csp_abdomens$age_cat_baseline=="<5 years"),]
cat2 = csp_abdomens[which(csp_abdomens$age_cat_baseline=="5-15 years"),]
cat3 = csp_abdomens[which(csp_abdomens$age_cat_baseline==">15 years"),]
cat1_mean = mean(cat1$haps_shared, na.rm = TRUE) 
cat2_mean = mean(cat2$haps_shared, na.rm = TRUE) 
cat3_mean = mean(cat3$haps_shared, na.rm = TRUE) 
pooled_sd = sd(append(cat2$haps_shared, cat3$haps_shared), na.rm = TRUE)
(cat2_mean - cat3_mean)/pooled_sd

# comparing haps_shared vs. village_name
kinesamo = csp_abdomens[which(csp_abdomens$village_name=="Kinesamo"),]
maruti = csp_abdomens[which(csp_abdomens$village_name=="Maruti"),]
kinesamo_mean = mean(kinesamo$haps_shared, na.rm = TRUE) 
maruti_mean = mean(maruti$haps_shared, na.rm = TRUE) 
pooled_sd = sd(append(kinesamo$haps_shared, maruti$haps_shared), na.rm = TRUE)
(kinesamo_mean - maruti_mean)/pooled_sd

# comparing haps_shared vs. pfr364Q_std_combined_cat
cat1 = csp_abdomens[which(csp_abdomens$pfr364Q_std_combined_cat=="over_rdt_detection"),]
cat2 = csp_abdomens[which(csp_abdomens$pfr364Q_std_combined_cat=="under_rdt_detection"),]
cat1_mean = mean(cat1$haps_shared, na.rm = TRUE) 
cat2_mean = mean(cat2$haps_shared, na.rm = TRUE) 
pooled_sd = sd(append(cat2$haps_shared, cat1$haps_shared), na.rm = TRUE)
(cat2_mean - cat1_mean)/pooled_sd

# comparing age_cat_baseline vs. village_name
table(csp_abdomens$village_name,csp_abdomens$age_cat_baseline, useNA = "always")
firstmodel = glm(village_name ~ age_cat_baseline, data=csp_abdomens, family=binomial(link="logit"))
summary(firstmodel)
exp(-0.0641)
exp(-0.7809)

# comparing age_cat_baseline vs. pfr364Q_std_combined_cat
table(csp_abdomens$pfr364Q_std_combined_cat,csp_abdomens$age_cat_baseline, useNA = "always")
firstmodel = glm(pfr364Q_std_combined_cat ~ age_cat_baseline, data=csp_abdomens, family=binomial(link="logit"))
summary(firstmodel)
exp(-1.3252)
exp(-1.7151)

# comparing village_name vs. pfr364Q_std_combined
table(csp_abdomens$village_name,csp_abdomens$pfr364Q_std_combined_cat, useNA = "always")
firstmodel = glm(pfr364Q_std_combined_cat ~ village_name, data=csp_abdomens, family=binomial(link="logit"))
summary(firstmodel)
exp(1.2826)

```


## write out the final model data set for csp abdomen sharing
```{r}
# write_csv(csp_abdomens,"Desktop/spat21_csp_abdomen_sharing_final_model_data_31OCT2019.csv")
# write_rds(csp_abdomens,"Desktop/spat21_csp_abdomen_sharing_final_model_data_31OCT2019.rds")
```









